---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Data
For this project, we use three datasets that provide insights into different aspects of restaurant and food-related data.

## Description
### Restaurant Inspection Data

We utilize the [DOHMH New York City Restaurant Inspection dataset](https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-pn8j/about_data) from the NYC OpenData library. This dataset includes details on all sustained or pending violation citations from full or special program inspections conducted within three years prior to the most recent inspection. The data covers restaurants and college cafeterias that are actively operating as of the RECORD DATE (the date the data was pulled).

**Key points about this dataset**:
	- This is a huge data set containing about 260,000 observations of 27 variables. The data is updated daily according to the description of the data.
	- Each violation record is associated with an inspection, and establishments with multiple violations will have repeated fields for each violation.
	- Establishments are uniquely identified by their CAMIS (record ID) number.
	- Restaurants that have started operations or gone out of business within the dataset period are included only if they were active on the RECORD DATE.
	- Records also include restaurants applying for permits but not yet inspected and inspections resulting in no violations.
	- Restaurants with an inspection date of “1/1/1900” are new establishments awaiting inspection.
	- Restaurants with no violations are represented by a single row coded as having no violations in the ACTION field.

### Fast Food Chain Analysis Data

For the analysis of the most popular fast food chains, we use a dataset that provides nationwide data for the United States, not limited to New York. This dataset is sourced from [Passby Blog’s Top Fast Food Chains Foot Traffic](https://passby.com/blog/top-fast-food-chains-foot-traffic/), offering insights into consumer behavior and fast food chain popularity across the country.

We build this dataset manually.

### Customer Satisfaction Data

To explore customer satisfaction, we use data from the [American Customer Satisfaction Index (ACSI)](https://theacsi.org/industries/restaurant/fast-food-restaurants/). This dataset evaluates customer satisfaction levels across various restaurant types and food delivery services, providing valuable insights into consumer preferences and trends.

We build this dataset manually.

## Missing value analysis 
### Restaurant Inspection Data

The dataset processing steps include the following:
	1.	Removing records with missing values in key features, such as restaurant names, inspection dates, and grades.
	2.	Converting categorical variables into factors to facilitate visualization.
	3.	(Optional) Creating additional features to enhance the analysis.

First, read the data:
```{r}
#library(tidyverse)
#library(ggrepel)
restaurant<-read_csv("/Users/zhiyizhang/Downloads/DOHMH_New_York_City_Restaurant_Inspection_Results_20241120.csv")
```

Get a general understanding of the data:
```{r}
#glimpse(restaurant)
summary(restaurant)
colnames(restaurant)
```

Check missing data, and remove key feature missing including restaurant names, inspection date, and grade.
```{r}
restaurant |> summarise(across(everything(), ~ sum(is.na(.))))

restaurant_clean <- restaurant |>
  drop_na(DBA, `INSPECTION DATE`, GRADE)
```

Detect grade distribution:
```{r}
grade_distribution <- restaurant_clean |>
  count(GRADE, sort = TRUE)
#grade_distribution
restaurant_clean |>
  ggplot(aes(x = GRADE)) +
  geom_bar() +
  labs(
    title = "Distribution of Grades",
    x = "Grade",
    y = "Count"
  ) +
  theme_minimal()
```

### Fast Food Chain Analysis Data
For this dataset, there is no need to address missing data. Instead, let’s focus on gaining a general understanding of the dataset.
```{r}
fast_food<-read_csv("/Users/zhiyizhang/Downloads/fast_food_chains_foot_traffic.csv")
colnames(fast_food)
```
Show the number of different fast food locations:
```{r}
fast_food |>
  ggplot(aes(x = reorder(Chain, USLocations), y = USLocations)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Number of U.S. Locations of Fast Food Chains",
    x = "Fast Food Chain",
    y = "Number of U.S. Locations"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Show the relationship between the number of US locations and global revenue:
```{r}
fast_food |>
  ggplot(aes(x = USLocations, y = `2023 Global Revenue (Billion USD)`, label = Chain)) +
  geom_point(aes(color = `Customer Satisfaction (%)`, size = `2023 Global Revenue (Billion USD)`)) + # Add size and color mapping
  geom_text_repel() + # Add labels with text repelling
  scale_x_log10() + # Log scale for the X-axis
  scale_y_log10() + # Log scale for the Y-axis
  labs(
    title = "Fast Food Chains: U.S. Locations vs. Global Revenue",
    x = "Total Number of U.S. Locations (Log10 Scale)",
    y = "2023 Global Revenue (Billion USD, Log10 Scale)",
    color = "Customer Satisfaction (%)",
    size = "2023 Global Revenue (Billion USD)"
  ) +
  theme_minimal()
```

### Customer Satisfaction Data
For this dataset, there is also no need to address missing data. Instead, let’s focus on gaining a general understanding of the dataset.
```{r}
customer_satisfaction<-read_csv("/Users/zhiyizhang/Downloads/customer_satisfation_2023_2024.csv")
colnames(customer_satisfaction)
```
Using line chart to show the satisfaction change of each company.
```{r}
customer_satisfaction <- customer_satisfaction |>
  mutate(across(`2023`:`2024`, as.numeric))
customer_satisfaction_long <- customer_satisfaction |>
  pivot_longer(cols = c(`2023`, `2024`), names_to = "Year", values_to = "Score") |>
  mutate(Year = as.numeric(Year))  # Convert Year to numeric

customer_satisfaction_long |>
  ggplot(aes(x = Year, y = Score, group = Company, color = Company)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = c(2023, 2024), labels = c("2023", "2024")) +
  labs(
    title = "Fast Food Company Trends: 2023 vs 2024",
    x = "Year",
    y = "Score",
    color = "Company"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

